Controlling synchronization in coupled area-preserving maps by stickiness

### 2018-09-26 Syncarea_control/paper.tex
paper updated

\title{Mechanism to control synchronization in coupled area-preserving maps}

@@

\begin{abstract}
 A unidirectionally coupled area-preserving maps with mixed phase space may 
show identical synchronization in the sticky neighborhood of the regular 
islands. We use this fact to devise a couple of numerical procedures to 
delay and expedite the process of synchronization in standard maps coupled 
under  Pecora-Caroll coupling scheme. The delay method is based on controlled 
kicking of trajectories away from synchronization traps for as long as 
necessary. On the other hand, the method to expiate the process is
achieved by a parameter perturbation technique which rapidly drives the 
chaotic trajectories to synchronization traps in the sticky neighborhoods 
of regular islands. We also point out of the limitations of these methods.
\end{abstract}

@@

\section{Introduction} 
Low-dimensional Hamiltonian systems commonly exhibit mixed phase space i.e. 
regular structures and chaotic regions may co-exist at a given degree of 
nonlinearity. This mixed nature has interesting consequences for the transport
properties of such systems, for instance, the existence of anomalous kinetics,
L\'{e}vy processes and L\'{e}vy flights \cite{Klafter1994,Zaburdaev2015}, 
power law contributions to recurrence and other statistics, and the existence 
of dynamical traps \cite{Zaslavsky2002a,Zaslavsky2002b}. An interesting region 
of a mixed phase space is the interface between regular regions and chaotic 
sea \cite{Mackay1984, Easton1993}. The dynamics in these neighbourhoods are 
complex but fairly well understood \cite{Meiss2015} for lower dimensional 
systems. The complexity arises from the fact that a chaotic trajectory spends 
an arbitrary long but finite time at the boundaries of regular islands before 
exiting to the chaotic sea. The intermittent tendency of chaotic trajectories 
to stay close to the regular boundaries is called the phenomenon of 
stickiness. A major consequence of the stickiness is  the existence of power 
law in the Poincar{\'e} recurrence times indicating algebraic decay for long 
times times rather than exponential decay expected for normal transport. 
Therefore, due to stickiness, even a small regular island can influence the 
global transport properties of the system and decay of correlations. The 
phenomenon has been of great interest and continues to be studied on 
theoretical level \cite{Altmann2006, Altmann2010,Livorati2012,Bunimovich2012, 
Kruger2015}. In addition, stickiness has found application in several problems 
such as particle advection in fluids \cite{Babiano1994,Tel2005}, transport in 
plasma fusion devices \cite{Szezech2012,Martins2014}, celestial mechanics 
\cite{Efthymiopoulos1999, Harsoula2010, Harsoula2016}.

A recent feature of stickiness has emerged from a connected quarter. In our 
recent work~\cite{Mahata2016}, we have looked at the effects of the mixed 
phase space on the synchronization in a system of two standard maps coupled in 
unidirectional drive-response configuration. We have shown that 
synchronization of chaotic trajectories of the drive and response maps 
typically occurs in the neighborhood of regular inlands as a consequence of 
stickiness in the region.  This is the first instance, as far as we are aware, 
where stickiness have been found to influence synchronization in coupled 
chaotic systems.  For a chaotic orbit, synchronization typically happens via 
an intermittent behavior in the phase difference of the drive and response 
maps.   The sticky neighborhoods of a regular islands contain dynamical traps 
in chaotic orbits. In such trapping regions in the phase space, parts of a 
chaotic trajectory are almost regular in time and allow for synchronization to 
occur. Such traps are characterized using the properties of the finite-time 
Lyapunov exponent \cite{Szezech2005}. In the context of our work, we will 
refer to these traps as synchronization traps. We further noted that the 
behavior of the synchronization traps can be analyzed in a more quantitative 
way by analyzing the location and stability properties of the periodic orbits 
in the vicinity of the locations where synchronization takes place.

Synchronization in coupled chaotic systems, coupled map lattices and networks 
have been studied in great details over the years. Coupled Hamiltonian systems are also known to show synchronization such as measure synchronization~\cite{Hampton1999,Wang2003,Vincent2005,Gupta2017} and identical synchronization~\cite{Mahata2016,Das2017} The fact however remains 
that the ability to control synchronization based on the dynamics of the 
process has not been given much attention, until 
recently~\cite{Grabow2011,Wang2016}. In realistic systems, the speed of 
synchronization may be significant. For example, in
neurosciences,  the speed of the visual and olfactory processing are 
interesting problems \cite{Thorpe1996,Uchida2003}. In this paper, we intend 
to develop a couple of 
mechanisms to control synchronization process to wit,  to increase and to 
decrease synchronization times based on the location of synchronization traps 
in the sticky neighborhood of regular islands in the phase space of the 
area-preserving system of the standard map. We demonstrate the mechanism at $K 
= 6$ where only two small regular islands exist. It is important to point out 
that the role of stickiness in Hamiltonian systems for synchronization was 
already predicted by Zaslavsky~\cite{Zaslavsky2002b} early in the last decade.

The rest of the paper is 
organized in the following way: we explain the coupling scheme in 
sec.~\ref{sec:Pecora_Carroll} and the mechanism based on the location of 
synchronization traps is given in sec.~\ref{Master stability function}. The 
numerical procedures to delay and advanced synchronization times are 
demonstrated in sec.~\ref{sec:delay} and sec.~\ref{sec:advanced} respectively, 
and the paper ends with conclusions in sec.~\ref{sec:conclusions} including 
discussion on limitations and implications.

@@

\begin{empheq}[right=\empheqrbrace \mod 1]{align}
P_{n+1} = P_n +  \frac{K}{2\pi}\sin(2\pi Q_n) \nonumber\\
Q_{n+1} = P_{n+1} + Q_n 
\end{empheq}

\begin{figure}[b]
	\includegraphics[scale=.45]{Standard_Map}
	\caption{\label{fig:Standard_map} \footnotesize The phase space of the standard map for  random initial conditions from the uniform distribution on $[0,1]$,  at the parameter value K = 6.0.}
\end{figure}
Here the subscript $n$ denotes the discrete time and $K$ is the nonlinearity parameter. These equations typically describe the evolution of two canonical variables $P$ and $Q$ which correspond to the momentum and co-ordinate in the Poincar\'{e} section of a freely moving rotator with interleaved periodic kicks. This system represents the behavior of a variety of systems such as charged particle confinement in  mirror magnetic traps, particle dynamics in accelerator, comet dynamics in solar systems etc.  Two-dimensional  phase space plots  of the standard map for the parameter value $K = 6$ using $25$ initial conditions  are shown in Figs.~\ref{fig:Standard_map}.

@@

\begin{minipage}[t]{0.5\textwidth}
	\begin{empheq}[right=\empheqrbrace \mod 1]{align}\nonumber
	\label{equ:drive}
	P^d_{n+1} = P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n)\nonumber\\
	Q^d_{n+1} = P^d_{n+1} + Q^d_n \nonumber
	\end{empheq}
	\centering \textbf{(Drive)}

\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
	\begin{empheq}[right=\empheqrbrace \mod 1]{align}
	P^r_{n+1} = P^d_n + \frac{K}{2\pi}\sin(2\pi Q^r_n) \nonumber \\
	Q^r_{n+1} = P^r_{n+1} + Q^r_n \nonumber
	\end{empheq}
	\centering \textbf{(Response)}
\end{minipage}

@@

The initial values of $Q$ in the drive and response maps are chosen arbitrarily, whereas the $P$ values are identical. The system is said to reach complete synchronization when both the $Q$ values of the drive and response systems evolve asymptotically to identical values i.e. 
\begin{equation}
\lim_{t\rightarrow\infty}(Q^d_n - Q^r_n) = 0
\end{equation}

@@

In all of the computations reported in the work, two chaotic trajectories are considered to be synchronized to numerical accuracy, if the Euclidean distance between them is less than $10^{-5}$. An example is shown in Fig.~\ref{fig:sync_ex}. Here, we plot $\Delta Q = Q^d_n-Q^r_n$ for initial conditions $(P^d_0,Q^d_0,Q^r_0) = (0.569, 0.906,0.106)$ at $K = 6$. The synchronization time is found to be 570246.

@@

\begin{align}
\frac{d\overbar{X}_d}{dt} = \overbar{F}(\overbar{X}_d) \nonumber  &&
\frac{d\overbar{X_r}}{dt} = \overbar{F}(\overbar{X}_r) + \alpha E(\overbar{X}_d-\overbar{X}_r)
\end{align}

Here $\overbar{X}_d$ and $\overbar{X}_r$ are drive and response variables; the matrix $E$ determines the linear combination of the $\overbar{X}$ used in the difference and $\alpha$ is the coupling strength. For the map case, we have the following form

\begin{align}
\overbar{X}^d_{n+1} = \overbar{F}(\overbar{X}^d_{n}) \nonumber  &&
\overbar{X}^r_{n+1} = \overbar{F}(\overbar{X}^r_{n}) + \alpha E(\overbar{X}^d_{n}-\overbar{X}^r_n)
\end{align}

Therefore, in the case of a general unidirectional coupling of two standard maps, we get
\begin{align}
P^d_{n+1} &= P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n) \nonumber\\
Q^d_{n+1} &= P^d_{n+1} + Q^d_n \nonumber\\
P^r_{n+1} &= P^r_n + \frac{K}{2\pi}\sin (2\pi Q^r_n) + \alpha(P^d_n-P^r_n) \nonumber \\
Q^r_{n+1} &= P^r_{n+1} + Q^r_n 
\label{equ:coupled}
\end{align}

@@

To find the stability of the synchronous state, we first express  Eq.(\ref{equ:coupled}) in terms of  $P^\perp = P^d - P^r$ and $Q^\perp = Q^d - Q^r$, as follows
\begin{eqnarray}
\label{equ:trans}
P^\perp_{n+1} = (1-\alpha)P^\perp_n+ \frac{K}{2\pi} \sin(2\pi Q^d_n) - \frac{K}{2\pi} \sin(2\pi Q^r_n) \\ \nonumber
Q^\perp_{n+1} = (1-\alpha)P^\perp_n+ Q^\perp_n + \frac{K}{2\pi} \sin(2\pi Q^d_n) - \frac{K}{2\pi} \sin(2\pi Q^r_n)  
\end{eqnarray}

We now write the variational equation for Eq.(\ref{equ:trans}) by linearizing about $(P^d_n,Q^d_n)$
\begin{equation}
\label{equ:variational}
\left[ \begin{array}{c} \delta P^\perp_{n+1} \\ \delta Q^\perp_{n+1} \end{array} \right] = \mathcal{M}(\alpha)\left[ \begin{array}{c}\delta P^\perp_{n} \\ \delta Q^\perp_{n} \end{array} \right]
\end{equation}
where the matrix $\mathcal{M}(\alpha)$ is given by
\begin{equation} 
\label{equ:M}
\mathcal{M}(\alpha) = \begin{bmatrix} 1-\alpha & K\cos(2\pi Q^d_n) \\ 1 - \alpha & 1 + K\cos(2\pi Q^d_n) \end{bmatrix} 
\end{equation}

This is the master stability equation for the unidirectionally coupled standard map. The variational equation (\ref{equ:variational}) is the master stability equation for the coupled system under investigation. The associated largest LE computed from the master stability equation is the master stability function of the system, given by:
\begin{equation}
\label{equ:msf}
\lambda  = \lim_{n \rightarrow \infty} \lim_{\delta\overbar{X}_0\rightarrow 0}\frac{1}{n}\sum^{n-1}_{i=0}\ln|JM^n(\overbar{X}_i)|
\end{equation}

@@

\begin{equation} 
\mathcal{M}(1)= \begin{bmatrix} 0 & K\cos(2\pi Q^d_n) \\ 0 & 1 + K\cos(2\pi Q^d_n) \end{bmatrix} 
\end{equation}

@@

\begin{equation}
\label{equ:LE}
\lambda   = \lim_{n \rightarrow \infty} \frac{1}{n}\sum^{n-1}_{i=0}\ln|1+K\cos(2\pi Q^d_n)|
\end{equation}

In general, we define the $kth$ time-$n$ Lyapunov exponent associated with an initial point $\overbar{X}_0 = (P_0,Q_0)$ for a map $M(P,Q)$ as
\begin{equation}
\label{equ:FLEf}
\lambda _k(\overbar{X}_0;n)  = \frac{1}{n}\sum^{n-1}_{i=0}\ln|JM^n(\overbar{X}_i)|
\end{equation}

@@

Here $n$ is a positive integer and $JM^n(\overbar{X}_i)$ denotes the Jacobian matrix of the $n$-times iterated map. We extend this notion to the master stability function defined in the Sec.~\ref{Master stability function}  i.e.  the non-zero Lyapunov exponent defined in Eq.(\ref{equ:LE}) 
which takes the following finite time version 
\begin{equation}
\lambda_1^s(\overbar{X}_0;n) = \frac{1}{n}\sum^{n-1}_{i=0}\ln|1 + K \cos (2\pi Q_i)|
\end{equation}

The subscript ($k = 1$) has been dropped hereafter as we have only one LE to compute.

We plot the variation in the time-5 FTLE defined above at $K = 6$. At the point of synchronization, the FTLE values attain a set of negative values consistently in a small window, indicating the existence of a synchronization trap.  In Fig.~\ref{fig:FTLE} , the three plots show the fluctuations  of FTLE near the point of synchronization for three different sets of initial conditions - $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a),  $(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c) with synchronization times 287090, 234521, and 164782 respectively. The values on the x-axis indicate time steps at which averages are computed so that synchronization times and the temporal neighborhood are effectively captured in the plots wherein a temporary drop is clearly visible, indicated by ellipses in red. Synchronization of chaotic trajectories typically occurs in one of such traps.

\section{Delayed Synchronization Times}
\label{sec:delay}
In order to develop a numerical technique to delay synchronization, we  first 
have to identify the vicinity of regular islands in the phase space. A 
mechanism  to suppress stickiness based on the knowledge of hyperbolic and non 
hyperbolic regions in the phase has been reported recently~\cite{Kruger2015}. 
Our method, however, targets the trajectories themselves in the sticky region 
to delay the process.  It is to be noted that the proposed procedure ignores 
the details of a rather complex hierarchy of structures around regular 
islands. We employ the edge-detection algorithm due to Benkadda {\it et al.} 
\cite{Benkadda1997} to identify the edges of regular islands in phase space. 
For the numerical procedure, we divide the phase space in $100 \times 100$ 
grid. The edges of both the islands are detected by applying the standard map 
to point initiated in the chaotic sea, for $10^7$ times.  The edges thus 
detected in shown in Fig~\ref{fig:edge_reg_island}.  We have chosen the 
thickness parameter to be $d=0.02$ i.e. euclidean distance of $d$ from points 
on the edge indicate the extent of the vicinity of the island.  This vicinity, 
therefore, indicate the domain wherein the synchronization traps exist.  To 
demonstrate this explicitly, we compare the phase angles, defined by $\theta = 
\tan^{-1}(\frac{Q}{P}$), of the points on the numerically detected edges and 
the points of synchronization, as 
follows.

@@

The proximity parameter $d_0 = 0.02$ defines the region in the neighborhood of 
the numerically detected edges wherein synchronization traps exist. The basic 
idea to control synchronization time is to identify the step at which the 
chaotic trajectory visits the numerically determined sticky neighborhood 
followed by a slight deflection so that the trajectory restarts at a point 
outside.  Our procedure depends upon how many times we deflect the trajectory  
which will be referred to as step control. This means that if an $n-$step 
control is  employed then, the trajectory will be kicked away from the domain 
$n$-times during its first $n$ visits i.e. once per visit. The deflection is 
achieved by adding a randomly generated fraction below $0.1$ to the point 
visiting the domain, and therefore the maximum deflection area is roughly 
$1\%$ of the phase space.  This procedure applied on a given set of initial 
conditions of the drive and response maps, may result in four possibilities of 
synchronization time -- (1) successful delay  \textbf{(S)}, (2) no delay 
\textbf{(N)}, (3) failed to achieve synchronization \textbf{(F)}, and (4) 
undesirable faster synchronization \textbf{(U)}. Clearly, the only first 
possibility is the aim of the mechanism; the last one constitute a hazard 
while the other two are trivial. 
\begin{figure}[h]
	\includegraphics[scale=0.45]{S_U_Percent}
	\caption{\label{fig:Control_success}\footnotesize The curves indicate the 
		percentages of (S) successful delays shown as the blue curve, and (U) 
		undesirable fast synchronization shown as the red curve. }
\end{figure}

For the numerical implementation of the mechanism, we have considered $1000$ randomly chosen initial conditions from the chaotic sea of the phase space at $K = 6$. The maps have been iterated for $10^7$ times in each case.  The sticky domain is the determined by the proximity parameter $d_0$ around the numerically detected edges of the regular islands. We demonstrate the effectiveness of the control procedure  with respect to several $n-$step controls. The computations have been performed for $n \in \mathbb{A}$, where $\mathbb{A} = \{10, 20,30,...,90, 100, 150, ... ,10 000\}$.

\begin{figure}[t]
	\includegraphics[scale=0.45]{Sync_time_dist.eps}
	\caption{\label{fig:Sync_time_dist}\footnotesize Distribution of 
		synchronization times without n-step control (the blue curve) and with 
		n-step 
		control (the red curve). The control delays synchronization times. }
\end{figure}

>>>>

\title{Controlling synchronization in coupled area-preserving maps by 
stickiness}

@@

\begin{abstract}
Unidirectionally coupled area-preserving maps with a mixed phase space may 
show identical synchronization in the sticky neighborhood of the regular 
islands. We use this fact to devise numerical procedures to 
delay and expedite the process of synchronization in two standard maps coupled 
under  the Pecora-Caroll coupling scheme. The delay method is based on 
controlled 
kicking of trajectories away from synchronization traps for as long as 
necessary. The method to expiate the process is
achieved by a parameter perturbation technique which rapidly drives the 
chaotic trajectories to synchronization traps in the sticky neighborhoods 
of regular islands. We also discuss the limitations of these methods.
\end{abstract}

@@

\section{Introduction} 
A low-dimensional Hamiltonian system commonly exhibits a mixed phase space 
i.e. regular structures and chaotic regions may co-exist at a given degree of 
nonlinearity. This mixed nature has interesting consequences for the transport
properties of such systems, for instance, the existence of anomalous kinetics,
L\'{e}vy processes and L\'{e}vy flights \cite{Klafter1994,Zaburdaev2015}, 
power law contributions to recurrence and other statistics, and the existence 
of dynamical traps \cite{Zaslavsky2002a,Zaslavsky2002b}. An interesting region 
of a mixed phase space is the interface between regular regions and chaotic 
sea \cite{Mackay1984, Easton1993}. The dynamics in these neighborhoods are 
complex but fairly well understood \cite{Meiss2015} for low dimensional 
systems. The complexity arises from the fact that a chaotic trajectory spends 
an arbitrary long but finite time at the boundaries of regular islands before 
exiting to the chaotic sea. The intermittent tendency of chaotic trajectories 
to stay close to the regular boundaries is called stickiness. A major 
consequence of the stickiness is  the existence of power law in the 
Poincar{\'e} recurrence times indicating algebraic decay for long 
times times rather than exponential decay expected for normal transport. 
Therefore, due to stickiness, even a small regular island can influence the 
global transport properties of the system and decay of correlations. The 
phenomenon has been of great interest and continues to be studied on 
theoretical level \cite{Altmann2006, Altmann2010,Livorati2012,Bunimovich2012, 
Kruger2015, Lange2016}. In addition, stickiness has found application in 
several problems such as particle advection in fluids 
\cite{Babiano1994,Tel2005}, transport in 
plasma fusion devices \cite{Szezech2012,Martins2014}, celestial mechanics 
\cite{Efthymiopoulos1999, Harsoula2010, Harsoula2016}.

A surprising feature of stickiness has emerged in our earlier 
work~\cite{Mahata2016}, we have looked at the effects of the mixed phase space 
on the synchronization in a system of two standard maps coupled in 
unidirectional drive-response configuration. We have shown that 
synchronization of chaotic trajectories of the drive and response maps 
typically occurs in the neighborhood of regular inlands as a consequence of 
stickiness in the region.  This is the first instance, as far as we are aware, 
where stickiness have been found to influence synchronization in coupled 
chaotic systems. It is important to point out that the role of stickiness in 
Hamiltonian systems for synchronization was already predicted by 
Zaslavsky~\cite{Zaslavsky2002b} early in the last decade.  For a chaotic 
orbit, synchronization typically happens via an intermittent behavior in the 
phase difference of the drive and response maps.   The sticky neighborhoods of 
a regular islands {\it temporarily} traps chaotic orbits. In such trapping 
regions in the phase space, parts of a  chaotic trajectory are almost regular 
in time and allow for synchronization to occur. Such traps are characterized 
using the properties of the finite-time Lyapunov exponent \cite{Szezech2005}. 
In the context of our work, we will refer to these traps as synchronization 
traps. We further note that the behavior of the synchronization traps can be 
analyzed in a more quantitative way by analyzing the location and stability 
properties of the periodic orbits in the vicinity of the locations where 
synchronization takes place.

Synchronization in coupled chaotic systems, coupled map lattices and networks 
have been studied in details (for reviews, see 
\cite{Pecora1998,Boccaletti2002, Pecora2015}). 
Coupled Hamiltonian systems are also known to show synchronization such as 
measure synchronization~\cite{Hampton1999,Wang2003,Vincent2005,Gupta2017} and 
identical synchronization~\cite{Mahata2016,Das2017} The fact however remains 
that the ability to control synchronization based on the dynamics of the 
process has not been given much attention, until 
recently~\cite{Grabow2011,Wang2016}. In realistic systems, the speed of 
synchronization may be significant. For example, in
neurosciences,  the speed of the visual and olfactory processing are 
interesting problems \cite{Thorpe1996,Uchida2003}. In this paper, we intend 
to develop a couple of 
mechanisms to control synchronization process to wit,  to increase and to 
decrease synchronization times based on the location of synchronization traps 
in the sticky neighborhood of regular islands in the phase space of the 
area-preserving system of the standard map. We demonstrate the mechanism at $K 
= 6$ where only two small regular islands exist.

The rest of the paper is organized in the following way: we explain the 
coupling scheme in sec.~\ref{sec:Pecora_Carroll} and the mechanism based on 
the location of synchronization traps is given in sec.~\ref{Master stability 
function}. The numerical procedures to delay and advance synchronization 
times are demonstrated in Sec.~\ref{sec:delay} and Sec.~\ref{sec:advanced} 
respectively, and the paper ends with conclusions in 
Sec.~\ref{sec:conclusions} including discussion on limitations and 
implications.

\begin{figure}[b]
    \includegraphics[scale=.45]{Standard_Map}
    \caption{\label{fig:Standard_map} \footnotesize The phase space of the 
    standard map for  random initial conditions from the uniform distribution 
    on $[0,1]$,  at the parameter value K = 6.0.}
\end{figure}

@@

\begin{empheq}[right=\empheqrbrace \mod 1 .]{align}
P_{n+1} = P_n +  \frac{K}{2\pi}\sin(2\pi Q_n) \nonumber\\
Q_{n+1} = P_{n+1} + Q_n 
\end{empheq}

\noindent Here the subscript $n$ denotes the discrete time and $K$ is the 
nonlinearity parameter. These equations typically describe the evolution of 
two canonical variables $P$ and $Q$ which correspond to the momentum and 
co-ordinate in the Poincar\'{e} section of a freely moving rotator with 
interleaved periodic kicks. This system represents the behavior of a variety 
of systems such as charged particle confinement in  mirror magnetic traps, 
particle dynamics in accelerator, comet dynamics in solar systems 
etc.~\cite{Chirikov1960,Izraelev1980,Chirikov1989,Chirikov2008}
Two-dimensional  phase space plots  of the standard map for the parameter 
value $K = 6$ using $25$ initial conditions  are shown in 
Figs.~\ref{fig:Standard_map}.

@@

\begin{minipage}[t]{0.5\textwidth}
	\begin{empheq}[right=\empheqrbrace \mod 1]{align}\nonumber
	\label{equ:drive}
	P^\textnormal{d}_{n+1} = P^\textnormal{d}_n + \frac{K}{2\pi}\sin(2\pi 
	Q^\textnormal{d}_n)\nonumber\\
	Q^\textnormal{d}_{n+1} = P^\textnormal{d}_{n+1} + Q^\textnormal{d}_n 
	\nonumber
	\end{empheq}
	\centering \textbf{(Drive)}

\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
	\begin{empheq}[right=\empheqrbrace \mod 1]{align}
	P^\textnormal{r}_{n+1} = P^\textnormal{d}_n + \frac{K}{2\pi}\sin(2\pi 
	Q^\textnormal{r}_n) \nonumber \\
	Q^\textnormal{r}_{n+1} = P^\textnormal{r}_{n+1} + Q^\textnormal{r}_n 
	\nonumber
	\end{empheq}
	\centering \textbf{(Response)}
\end{minipage}

@@

The initial values of $Q$ in the drive and response maps are chosen 
arbitrarily, whereas the $P$ values are identical. The system is said to reach 
complete synchronization when both the $Q$ values of the drive and response 
systems evolve identically i.e. 
\begin{equation}
\lim_{t\rightarrow\infty}(Q^\textnormal{d}_n - Q^\textnormal{r}_n) = 0
\end{equation}

@@

In all of the computations reported in the work, two chaotic trajectories are 
considered to be synchronized to numerical accuracy, if the Euclidean distance 
between them is less than $10^{-5}$. An example is shown in 
Fig.~\ref{fig:sync_ex}. Here, we plot $\Delta Q = 
Q^\textnormal{d}_n-Q^\textnormal{r}_n$ for initial 
conditions $(P^\textnormal{d}_0,Q^\textnormal{d}_0,Q^\textnormal{r}_0) = 
(0.569, 0.906,0.106)$ at $K = 6$. The synchronization time is found to be 5,70,
246.

@@

\begin{align}
\frac{d\overbar{X}_d}{dt} & = \overbar{F}(\overbar{X}_d) \nonumber  \\
\frac{d\overbar{X_r}}{dt} &= \overbar{F}(\overbar{X}_r) + \alpha 
E(\overbar{X}_d-\overbar{X}_r).
\end{align}

\noindent Here $\overbar{X}_d$ and $\overbar{X}_r$ are drive and response 
variables; the matrix $E$ determines the linear combination of the 
$\overbar{X}$ used in the difference and $\alpha$ is the coupling strength. 
For the map case, we have the following form

\begin{align}
\overbar{X}^d_{n+1} &= \overbar{F}(\overbar{X}^d_{n}) \nonumber  \\
\overbar{X}^r_{n+1} &= \overbar{F}(\overbar{X}^r_{n}) + \alpha 
E(\overbar{X}^d_{n}-\overbar{X}^r_n).
\end{align}

Therefore, in the case of a general unidirectional coupling of two standard maps, we get
\begin{align}
P^d_{n+1} &= P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n) \nonumber\\
Q^d_{n+1} &= P^d_{n+1} + Q^d_n \nonumber\\
P^r_{n+1} &= P^r_n + \frac{K}{2\pi}\sin (2\pi Q^r_n) + \alpha(P^d_n-P^r_n) \nonumber \\
Q^r_{n+1} &= P^r_{n+1} + Q^r_n.
\label{equ:coupled}
\end{align}

@@

To find the stability of the synchronous state, we first express  
Eq.(\ref{equ:coupled}) in terms of  $P^\perp = P^\textnormal{d} - 
P^\textnormal{r}$ and $Q^\perp = Q^\textnormal{d} -Q^\textnormal{r}$, as 
follows
\begin{eqnarray}
\label{equ:trans}
P^\perp_{n+1} &= (1-\alpha)P^\perp_n+ \frac{K}{2\pi} \sin(2\pi Q^d_n) - 
\frac{K}{2\pi} \sin(2\pi Q^r_n) \nonumber \\ 
Q^\perp_{n+1} &= (1-\alpha)P^\perp_n+ Q^\perp_n + \frac{K}{2\pi} \sin(2\pi 
Q^d_n) \nonumber \\ &- \frac{K}{2\pi} \sin(2\pi Q^r_n).
\end{eqnarray}

We now write the variational equation for Eq.(\ref{equ:trans}) by linearizing 
about $(P^\textnormal{d}_n,Q^\textnormal{d}_n)$
\begin{equation}
\label{equ:variational}
\left[ \begin{array}{c} \delta P^\perp_{n+1} \\ \delta Q^\perp_{n+1} 
\end{array} \right] = \mathcal{M}(\alpha)\left[ \begin{array}{c}\delta 
P^\perp_{n} \\ \delta Q^\perp_{n} \end{array} \right],
\end{equation}
where the matrix $\mathcal{M}(\alpha)$ is given by
\begin{equation} 
\label{equ:M}
\mathcal{M}(\alpha) = \begin{bmatrix} 1-\alpha & K\cos(2\pi Q^d_n) \\ 1 - 
\alpha & 1 + K\cos(2\pi Q^d_n) \end{bmatrix}.
\end{equation}

This is the master stability equation for the unidirectionally coupled 
standard map. The variational equation (\ref{equ:variational}) is the master 
stability equation for the coupled system under investigation. The associated 
largest Lyapunov exponent computed from the master stability equation is the 
master stability function of the system, given by:
\begin{equation}
\label{equ:msf}
\lambda  = \lim_{n \rightarrow \infty} \lim_{\delta\overbar{X}_0\rightarrow 
0}\frac{1}{n}\sum^{n-1}_{i=0}\ln|JM^n(\overbar{X}_i)|.
\end{equation}

@@

\begin{equation} 
\mathcal{M}(1)= \begin{bmatrix} 0 & K\cos(2\pi Q^d_n) \\ 0 & 1 + K\cos(2\pi 
Q^d_n) \end{bmatrix},
\end{equation}

@@

\begin{equation}
\label{equ:LE}
\lambda   = \lim_{n \rightarrow \infty} 
\frac{1}{n}\sum^{n-1}_{i=0}\ln|1+K\cos(2\pi Q^d_n)|.
\end{equation}

In general, we define the $kth$ time-$n$ Lyapunov exponent associated with an initial point $\overbar{X}_0 = (P_0,Q_0)$ for a map $M(P,Q)$ as
\begin{equation}
\label{equ:FLEf}
\lambda _k(\overbar{X}_0;n)  = 
\frac{1}{n}\sum^{n-1}_{i=0}\ln|JM^n(\overbar{X}_i)|.
\end{equation}

@@

Here $n$ is a positive integer and $JM^n(\overbar{X}_i)$ denotes the Jacobian matrix of the $n$-times iterated map. We extend this notion to the master stability function defined in the Sec.~\ref{Master stability function}  i.e.  the non-zero Lyapunov exponent defined in Eq.(\ref{equ:LE}) 
which takes the following finite time version 
\begin{equation}
\lambda_1^s(\overbar{X}_0;n) = \frac{1}{n}\sum^{n-1}_{i=0}\ln|1 + K \cos (2\pi 
Q_i)|.
\end{equation}

The subscript ($k = 1$) has been dropped hereafter as we have only one 
exponent to compute.

We plot the variation in the time-5 FTLE defined above at $K = 6$. At the 
point of synchronization, the FTLE values attain a set of smaller values 
consistently in a small window, indicating the existence of a synchronization 
trap.  In Fig.~\ref{fig:FTLE} , the three plots show the fluctuations  of FTLE 
near the point of synchronization for three different sets of initial 
conditions - $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a),  
$(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c) with 
synchronization times 287090, 234521, and 164782 respectively. The values on 
the x-axis indicate time steps at which averages are computed so that 
synchronization times and the temporal neighborhood are effectively captured 
in the plots wherein a temporary drop is clearly visible, indicated by 
ellipses in red. Synchronization of chaotic trajectories typically occurs in 
one of such traps.

\section{Delayed Synchronization Times}
\label{sec:delay}
In order to develop a numerical technique to delay synchronization, we  first 
have to identify the vicinity of regular islands in the phase space. A 
mechanism  to suppress stickiness based on the knowledge of hyperbolic and non 
hyperbolic regions in the phase has been reported recently~\cite{Kruger2015}. 
Our method, however, targets the trajectories themselves in the sticky region 
to delay the process.  It is to be noted that the proposed procedure ignores 
the details of a rather complex hierarchy of structures around regular 
islands. We employ the edge-detection algorithm due to Benkadda {\it et al.} 
\cite{Benkadda1997} to identify the edges of regular islands in phase space. 
For the numerical procedure, we divide the phase space in $100 \times 100$ 
grid. The edges of both islands are detected by applying the standard map 
to point initiated in the chaotic sea, for $10^7$ times.  The edges thus 
detected in shown in Fig~\ref{fig:edge_reg_island}.  We have chosen the 
proximity parameter $d=0.02$ i.e. Euclidean distance from points 
on the edge indicate the extent of the vicinity of the island.  This vicinity, 
therefore, indicate the domain wherein the synchronization traps exist.  To 
demonstrate this explicitly, we compare the phase angles, defined by $\theta = 
\tan^{-1}(\frac{Q}{P}$), of the points on numerically detected edges and 
the points of synchronization, as 
follows.

@@

The proximity parameter $d_0 = 0.02$ defines the region in the neighborhood of 
the numerically detected edges wherein synchronization traps exist. The basic 
idea to control synchronization time is to identify the step at which the 
chaotic trajectory visits the numerically determined sticky neighborhood 
followed by a slight deflection so that the trajectory restarts at a point 
outside.  Our procedure depends upon how many times we deflect the trajectory  
which will be referred to as step control. This means that if an $n-$step 
control is  employed then, the trajectory will be kicked away from the domain 
$n$-times during its first $n$ visits i.e. once per visit. The deflection is 
achieved by adding a randomly generated fraction below $0.1$ to the point 
visiting the domain, and therefore the maximum deflection area is roughly 
$1\%$ of the phase space.  This procedure applied on a given set of initial 
conditions of the drive and response maps, may result in four possibilities of 
synchronization time -- (1) successful delay  \textbf{(S)}, (2) no delay 
\textbf{(N)}, (3) failed to achieve synchronization \textbf{(F)}, and (4) 
undesirable faster synchronization \textbf{(U)}. Clearly, the only first 
possibility is the aim of the mechanism; the last one constitute a hazard 
while the other two are trivial. 
\begin{figure}[h]
	\includegraphics[scale=0.42]{S_U_Percent}
	\caption{\label{fig:Control_success}\footnotesize The curves indicate the 
		percentages of (S) successful delays shown as the blue curve, and (U) 
		undesirable fast synchronization shown as the red curve. }
\end{figure}

For the numerical implementation of the mechanism, we have considered $1000$ 
randomly chosen initial conditions from the chaotic sea of the phase space at 
$K = 6$. The maps have been iterated for $10^7$ times in each case.  The 
sticky domain is the determined by the proximity parameter $d_0$ around the 
numerically detected edges of the regular islands. We demonstrate the 
effectiveness of the control procedure  with respect to several $n-$step 
controls. The computations have been performed for $n \in \mathbb{A}$, where 
$\mathbb{A} = \{10, 20,30,...,90, 100, 150,200, ... ,10 000\}$.

\begin{figure}[t]
	\includegraphics[scale=0.42]{Sync_time_dist.eps}
	\caption{\label{fig:Sync_time_dist}\footnotesize Distribution of 
		synchronization times without n-step control (the blue curve) and with 
		n-step 
		control (the red curve). The control delays synchronization times. }
\end{figure}
